---
title: "Plotting of OTUs responding to amendment in the Incubation Project"
output:
  html_document:
    df_print: paged
    fig_width: 8.5
    fig_height: 11
    fig_caption: true
    toc: true
    toc_float: true
    number_sections: false
    code_folding: hide
---
```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("."))
```

```{r, echo = F, message = F, warning = F}
# Load libraries
library(tidyverse)
library(phyloseq)
library(DESeq2)
library(naniar)
library(ggrepel)
library(viridis)
library(knitr)
library(kableExtra)
```

```{r, message = F, warning = F}
# Load phyloseq object as inc.physeq, not rarefied but unclassified OTUs removed if at phylum level
inc.physeq <- readRDS("data/RDS/not.rare.nounclass")
source("code/functions.R")
# LFC calculation function
who_diff_day <- function(DDS, choice1, choice2, phy.object){
  res = results(DDS, contrast = c("Treatment_Response", choice1, choice2), cooksCutoff = FALSE)
  #plotCounts(DDS, gene="OTU_311", intgroup="day")
  #Use above line to check if an OTU is increasing or decreasing depending on order of contrast
  alpha = 0.01
  #alpha = 0.1
  sigtab = res[which(res$padj < alpha), ]
  sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(phy.object)[rownames(sigtab), ], "matrix"))
  theme_set(theme_bw())
  scale_fill_discrete <- function(palname = "Set1", ...) {
    scale_fill_brewer(palette = palname, ...)
  }
  # Phylum order
  x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
  x = sort(x, TRUE)
  sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))
  # Genus order
  x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
  x = sort(x, TRUE)
  sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))
  #ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=phylum)) + geom_point(size=2) + 
  #  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=1.0)) +
  #  ggtitle("Day 0 to Day 7")
  return(sigtab)
}

# function plot log2FoldChange 
log_plot <- function(sigtab,t1){
  sigtab <- sigtab %>%
    rownames_to_column(var = "OTU") %>%
    filter(log2FoldChange >= 4) 
  
  ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=Phylum)) + geom_point(size=2) + 
    coord_flip() +
    ggtitle(t1)
} 
```

# Alfalfa Responders   

## Early
```{r, message = F, warning = F}
## Early   
alf.early <- subset_samples(inc.physeq, Treatment_Response %in% c("Alfalfa_early", "Reference_early")) %>%
  filter_taxa(function(x) sum(x) >= 3, T) 

# Be very careful of the design formula in the who_diff_day() function
# This function also selects only LFC >= 2 and alpha 0.01 for significant and increasing otus to be returned
log.plot.early.alf <- alf.early %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("Alfalfa_early", "Reference_early", alf.early) %>%
  log_plot("Alfalfa OTUS in early group that are significantly changing compared to reference early")

#AlfalfaDDS <- alf.early %>%
#  phyloseq_to_deseq2( ~ Treatment_Response) %>%
#  DESeq(test = "Wald", fitType = "local") 

#plotCounts(AlfalfaDDS, gene = "Otu00002", intgroup = c("Treatment_Response"))
#Use above line to check if an OTU is increasing or decreasing depending on order of contrast, replace with OTU of interest, run the phyloseq to desey in #other areas to check compost or mix. 

# Save a data frame of these results
log.plot.early.alf.data <- log.plot.early.alf$data %>%
  mutate(trt = c("Alfalfa_early"))

# print plot with viridis color 
log.plot.early.alf + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
```



## Late    
```{r, message = F, warning = F}
## Late 
alf.late <- subset_samples(inc.physeq, Treatment_Response %in% c("Alfalfa_late", "Reference_late")) %>%
  filter_taxa(function(x) sum(x) >= 3, T)

# Make deseq and plot as above
log.plot.late.alf <- alf.late %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("Alfalfa_late", "Reference_late", alf.late) %>%
  log_plot("Alfalfa OTUS in late group that are significantly changing compared to reference late")

# Save a data frame of these results
log.plot.late.alf.data <- log.plot.late.alf$data %>%
  mutate(trt = c("Alfalfa_late"))

# print plot with viridis color 
log.plot.late.alf + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
```

```{r, message = F, warning = F}
# Common to both early and late with LFC >=2
otustokeep <- intersect(log.plot.early.alf.data$OTU, log.plot.late.alf.data$OTU)

# Trim and rename variables
early_alf_OTUS <- log.plot.early.alf.data %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Alfalfa_early_log2FoldChange = log2FoldChange) 

late_alf_OTUS <- log.plot.late.alf.data %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Alfalfa_late_log2FoldChange = log2FoldChange)

# join early and late
all_alf <- full_join(early_alf_OTUS, late_alf_OTUS)
```

## Plot Alfalfa OTUs with LFC >= 4 in both early and late  

This plot is showing the common OTUs with LFC > 4 while also showing the LFC of OTUs observed in only early or late, represented by the points landing below 4 on either axis. Common OTUs with LFC < 4 are left unlabeled.   

```{r, message = F, warning = F, out.width = "100%"}
p <- ggplot(all_alf,
       aes(x = Alfalfa_early_log2FoldChange, 
           y = Alfalfa_late_log2FoldChange,
           color = Phylum,
           label = OTU)) +
  geom_miss_point() +
  geom_text_repel(aes(label=ifelse(Alfalfa_early_log2FoldChange>4 & Alfalfa_late_log2FoldChange>4,as.character(OTU),'')),hjust=1,vjust=1) +
  theme(legend.position = "none")

p + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
# Save this list of alfalfa LFC > 2 for more plots
saveRDS(all_alf, file = "data/RDS/LFC_alf_OTUs_June19.RDS")

png("Figures/early_late_alf_responders.png",height=5,width=8,units='in',res=300)
p + scale_colour_viridis_d(option = "viridis") +
  theme_bw() +
  theme(legend.position = "none")
dev.off()

```

## Plot a table of the Alfalfa OTUs with taxonomy   

These OTUs had response greater than 4 in both early and late alfalfa   

```{r, message = F, warning = F}
OTUs <- all_alf %>%
  filter(Alfalfa_early_log2FoldChange > 4 & Alfalfa_late_log2FoldChange > 4)
write.table(OTUs, file = "Figures/alf_responders.txt", sep = ",", quote = F, row.names = F)
kable(OTUs) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "400px")
```

# WORK HERE

Let's visualize the relative abundance of the significantly responding genera in each treatment. We will use a sacked area chart.

https://www.r-graph-gallery.com/136-stacked-area-chart.html

```{r}
alf.phy <- subset_samples(inc.physeq, treatment %in% c("Alfalfa") & day %in% c("0", "7", "14", "21", "35", "49", "97")) %>%
  filter_taxa(function(x) sum(x) >= 3, T) 

rela.alf.phy <- merge_samples(alf.phy, "day") %>%
  transform_sample_counts(function(x) x / sum(x)) 

sample_data(rela.alf.phy)$day <- levels(sample_data(alf.phy)$day)

merge.phy <- rela.alf.phy %>%
  psmelt() 

merge.phy <- merge.phy %>%
  mutate(Responder = ifelse(OTU %in% OTUs$OTU, "Responding", "Not_responding")) %>%
  filter(Responder == "Responding")

data <- merge.phy %>%
  select(day, OTU, Abundance, Responder, Genus)

pp <- ggplot(data, aes(x= as.numeric(day), y=Abundance, fill = OTU)) + 
    geom_area(aes(fill = OTU), alpha=0.6 , size=1, colour="black")

pp + 
  scale_x_continuous(breaks = c(0, 7, 14, 21, 35, 49, 97)) +
  xlab("Day") +
  ylab("Relative abundance \n of responders") 
```

### Two Otus in the top 10 of alfalfa responders were from the Genus Pseudomonas     
BLASTn results for these two Otu00064 and Otu00494 were most similar to sequences from Pseudomonas fulva and Pseudomonas putida, both members of the Pseudomonas putida group

### Plot counts of interesting Otu00064 and Otu00494 early    
```{r, message = F, warning = F}
early <- alf.early %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") 
late <- alf.late %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local")
plotCounts(early, gene = c("Otu00064"), intgroup = c("Treatment_Response"))
plotCounts(early, gene = c("Otu00494"), intgroup = c("Treatment_Response"))
#Use above line to check if an OTU is increasing or decreasing depending on order of contrast, replace with OTU of interest, run the phyloseq to desey in #other areas to check compost or mix. 
```

### Plot counts of interesting Otu00064 and Otu00494 late     
```{r, message = F, warning = F}
plotCounts(late, gene = "Otu00064", intgroup = c("Treatment_Response"))
plotCounts(late, gene = "Otu00494", intgroup = c("Treatment_Response"))
```

# Compost Responders   

## Early
```{r, message = F, warning = F}
## Early   
comp.early <- subset_samples(inc.physeq, Treatment_Response %in% c("Compost_early", "Reference_early")) %>%
  filter_taxa(function(x) sum(x) >= 3, T) 

# Be very careful of the design formula in the who_diff_day() function
# This function also selects only LFC >= 2 and alpha 0.01 for significant and increasing otus to be returned
log.plot.early.comp <- comp.early %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("Compost_early", "Reference_early", comp.early) %>%
  log_plot("Compost OTUS in early group that are significantly changing compared to reference early")

# Save a data frame of these results
log.plot.early.comp.data <- log.plot.early.comp$data %>%
  mutate(trt = c("Compost_early"))

# print plot with viridis color 
log.plot.early.comp + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
```

## Late    
```{r, message = F, warning = F}
## Late 
comp.late <- subset_samples(inc.physeq, Treatment_Response %in% c("Compost_late", "Reference_late")) %>%
  filter_taxa(function(x) sum(x) >= 3, T)

# Make deseq and plot as above
log.plot.late.comp <- comp.late %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("Compost_late", "Reference_late", comp.late) %>%
  log_plot("Compost OTUS in late group that are significantly changing compared to reference late")

# Save a data frame of these results
log.plot.late.comp.data <- log.plot.late.comp$data %>%
  mutate(trt = c("Compost_late"))

# print plot with viridis color 
log.plot.late.comp + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
```

```{r, message = F, warning = F}
# Common to both early and late with LFC >=2
otustokeep <- intersect(log.plot.early.comp.data$OTU, log.plot.late.comp.data$OTU)

# Trim and rename variables
early_comp_OTUS <- log.plot.early.comp.data %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Compost_early_log2FoldChange = log2FoldChange) 

late_comp_OTUS <- log.plot.late.comp.data %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Compost_late_log2FoldChange = log2FoldChange)

# join early and late
all_comp <- full_join(early_comp_OTUS, late_comp_OTUS)
```

## Plot Compost OTUs with LFC >= 4 in both early and late  

This plot is showing the common OTUs with LFC > 4 while also showing the LFC of OTUs observed in only early or late, represented by the points landing below 4 on either axis. Common OTUs with LFC < 4 are left unlabeled.   

```{r, message = F, warning = F, out.width = "100%"}
p <- ggplot(all_comp,
       aes(x = Compost_early_log2FoldChange, 
           y = Compost_late_log2FoldChange,
           color = Phylum,
           label = OTU)) +
  geom_miss_point() +
  geom_text_repel(aes(label=ifelse(Compost_early_log2FoldChange>4 & Compost_late_log2FoldChange>4,as.character(OTU),'')),hjust=0,vjust=0) 

p + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
# Save this list of compcompa LFC > 2 for more plots
saveRDS(all_comp, file = "data/RDS/LFC_comp_OTUs_June19.RDS")
```

## Plot a table of the Compost OTUs with taxonomy   

These OTUs had response greater than 3 in both early and late compost  

```{r, message = F, warning = F}
OTUs <- all_comp %>%
  filter(Compost_early_log2FoldChange>4 & Compost_late_log2FoldChange>4)

kable(OTUs) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "400px")
```

# Mix Responders   

## Early
```{r, message = F, warning = F}
## Early   
mix.early <- subset_samples(inc.physeq, Treatment_Response %in% c("Mix_early", "Reference_early")) %>%
  filter_taxa(function(x) sum(x) >= 3, T) 

# Be very careful of the design formula in the who_diff_day() function
# This function also selects only LFC >= 2 and alpha 0.01 for significant and increasing otus to be returned
log.plot.early.mix <- mix.early %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("Mix_early", "Reference_early", mix.early) %>%
  log_plot("Mix OTUS in early group that are significantly changing mixared to reference early")

# Save a data frame of these results
log.plot.early.mix.data <- log.plot.early.mix$data %>%
  mutate(trt = c("Mix_early"))

# print plot with viridis color 
log.plot.early.mix + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
```

## Late    
```{r, message = F, warning = F}
## Late 
mix.late <- subset_samples(inc.physeq, Treatment_Response %in% c("Mix_late", "Reference_late")) %>%
  filter_taxa(function(x) sum(x) >= 3, T)

# Make deseq and plot as above
log.plot.late.mix <- mix.late %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("Mix_late", "Reference_late", mix.late) %>%
  log_plot("Mix OTUS in late group that are significantly changing mixared to reference late")

# Save a data frame of these results
log.plot.late.mix.data <- log.plot.late.mix$data %>%
  mutate(trt = c("Mix_late"))

# print plot with viridis color 
log.plot.late.mix + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
```

```{r, message = F, warning = F}
# Common to both early and late with LFC >=2
otustokeep <- intersect(log.plot.early.mix.data$OTU, log.plot.late.mix.data$OTU)

# Trim and rename variables
early_mix_OTUS <- log.plot.early.mix.data %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Mix_early_log2FoldChange = log2FoldChange) 

late_mix_OTUS <- log.plot.late.mix.data %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Mix_late_log2FoldChange = log2FoldChange)

# join early and late
all_mix <- full_join(early_mix_OTUS, late_mix_OTUS)
```

## Plot Mix OTUs with LFC >= 4 in both early and late  

This plot is showing the common OTUs with LFC > 4 while also showing the LFC of OTUs observed in only early or late, represented by the points landing below 4 on either axis. Common OTUs with LFC < 4 are left unlabeled.   

```{r, message = F, warning = F, out.width = "100%"}
p <- ggplot(all_mix,
       aes(x = Mix_early_log2FoldChange, 
           y = Mix_late_log2FoldChange,
           color = Phylum,
           label = OTU)) +
  geom_miss_point() +
  geom_text_repel(aes(label=ifelse(Mix_early_log2FoldChange>4 & Mix_late_log2FoldChange>4,as.character(OTU),'')),hjust=0,vjust=0) 

p + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
# Save this list of mixmixa LFC > 2 for more plots
saveRDS(all_mix, file = "data/RDS/LFC_mix_OTUs_June19.RDS")
```

## Plot a table of the Mix OTUs with taxonomy   

These OTUs had response greater than 4 in both early and late mix   

```{r, message = F, warning = F}
OTUs <- all_mix %>%
  filter(Mix_early_log2FoldChange>4 & Mix_late_log2FoldChange>4)

kable(OTUs) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "400px")
```

## Plot relative abundance of OTUs with LFC > 4
```{r}
# We have lists of LFC OTUs, need to create a function that takes this and a target response group and returns the average relative abundance for these
# input: list of OTUs, Treatment group
resp_alf <- readRDS("data/RDS/LFC_alf_OTUs_June19.RDS") %>% dplyr::rename(label = OTU)
OTUs <- as.character(resp_alf$label)
alf.rela.early <- RelaOTUs(inc.physeq, c("Alfalfa_early"), OTUs) %>%
  select(label = OTU, Alfalfa_early_meanRela = mean)
alf.rela.late <- RelaOTUs(inc.physeq, c("Alfalfa_late"), OTUs) %>%
  select(label = OTU, meanAlfalfa_late_meanRela = mean)
all_alf_rela <- full_join(resp_alf, alf.rela.early) %>%
  full_join(alf.rela.late)
```

