---
title: "Plotting of OTUs responding to amendment in the Incubation Project"
output:
  html_document:
    df_print: paged
    fig_width: 8.5
    fig_height: 11
    fig_caption: true
    toc: true
    number_sections: false
---
```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("."))
```

#### Loading needed libraries, check if you need to install packages
```{r, echo = F, message = F, warning = F}
# Load libraries
library(tidyverse)
library(phyloseq)
library(DESeq2)
library(naniar)
library(ggrepel)
library(viridis)
library(knitr)
library(kableExtra)
```

#### Read in phyloseq object and custom deseq functions
```{r, message = F, warning = F}
# Load phyloseq object as inc.physeq, not rarefied but unclassified OTUs removed if at phylum level
inc.physeq <- readRDS("data/RDS/not.rare.nounclass")

# LFC calculation function
who_diff_day <- function(DDS, choice1, choice2, phy.object){
  res = results(DDS, contrast = c("Treatment_Response", choice1, choice2), cooksCutoff = FALSE)
  #plotCounts(AlfalfaDDS, gene="OTU_311", intgroup="day")
  #Use above line to check if an OTU is increasing or decreasing depending on order of contrast
  alpha = 0.01
  #alpha = 0.1
  sigtab = res[which(res$padj < alpha), ]
  sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(phy.object)[rownames(sigtab), ], "matrix"))
  theme_set(theme_bw())
  scale_fill_discrete <- function(palname = "Set1", ...) {
    scale_fill_brewer(palette = palname, ...)
  }
  # Phylum order
  x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
  x = sort(x, TRUE)
  sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))
  # Genus order
  x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
  x = sort(x, TRUE)
  sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))
  #ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=phylum)) + geom_point(size=2) + 
  #  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=1.0)) +
  #  ggtitle("Day 0 to Day 7")
  return(sigtab)
}

# function plot log2FoldChange 
log_plot <- function(sigtab,t1){
  sigtab <- sigtab %>%
    rownames_to_column(var = "OTU") %>%
    filter(log2FoldChange >= 2) 
  
  ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=Phylum)) + geom_point(size=2) + 
    coord_flip() +
    ggtitle(t1)
} 
```

# Alfalfa Responders   

## Early
```{r, message = F, warning = F}
## Early   
alf.early <- subset_samples(inc.physeq, Treatment_Response %in% c("Alfalfa_early", "Reference_early")) %>%
  filter_taxa(function(x) sum(x) >= 3, T) 

# Be very careful of the design formula in the who_diff_day() function
# This function also selects only LFC >= 2 and alpha 0.01 for significant and increasing otus to be returned
log.plot.early.alf <- alf.early %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("Alfalfa_early", "Reference_early", alf.early) %>%
  log_plot("Alfalfa OTUS in early group that are significantly changing compared to reference early")

# Save a data frame of these results
log.plot.early.alf.data <- log.plot.early.alf$data %>%
  mutate(trt = c("Alfalfa_early"))

# print plot with viridis color 
log.plot.early.alf + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
```

## Late    
```{r, message = F, warning = F}
## Late 
alf.late <- subset_samples(inc.physeq, Treatment_Response %in% c("Alfalfa_late", "Reference_late")) %>%
  filter_taxa(function(x) sum(x) >= 3, T)

# Make deseq and plot as above
log.plot.late.alf <- alf.late %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("Alfalfa_late", "Reference_late", alf.late) %>%
  log_plot("Alfalfa OTUS in late group that are significantly changing compared to reference late")

# Save a data frame of these results
log.plot.late.alf.data <- log.plot.late.alf$data %>%
  mutate(trt = c("Alfalfa_late"))

# print plot with viridis color 
log.plot.late.alf + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
```

```{r, message = F, warning = F}
# Common to both early and late with LFC >=2
otustokeep <- intersect(log.plot.early.alf.data$OTU, log.plot.late.alf.data$OTU)

# Trim and rename variables
early_alf_OTUS <- log.plot.early.alf.data %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Alfalfa_early_log2FoldChange = log2FoldChange) 

late_alf_OTUS <- log.plot.late.alf.data %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Alfalfa_late_log2FoldChange = log2FoldChange)

# join early and late
all_alf <- full_join(early_alf_OTUS, late_alf_OTUS)
```

## Plot Alfalfa OTUs with LFC >= 2 in both early and late  

This plot is showing the common OTUs with LFC > 3 while also showing the LFC of OTUs observed in only early or late, represented by the points landing below 2 on either axis. Common OTUs with LFC < 3 are left unlabeled.   

```{r, message = F, warning = F, out.width = "100%"}
p <- ggplot(all_alf,
       aes(x = Alfalfa_early_log2FoldChange, 
           y = Alfalfa_late_log2FoldChange,
           color = Phylum,
           label = OTU)) +
  geom_miss_point() +
  geom_text_repel(aes(label=ifelse(Alfalfa_early_log2FoldChange>3 & Alfalfa_late_log2FoldChange>3,as.character(OTU),'')),hjust=0,vjust=0) 

p + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
# Save this list of alfalfa LFC > 2 for more plots
saveRDS(all_alf, file = "data/RDS/LFC_alf_OTUs_June19.RDS")
```

## Plot a table of the Alfalfa OTUs with taxonomy   

These OTUs had response greater than 3 in both early and late alfalfa   

```{r, message = F, warning = F}
OTUs <- all_alf %>%
  filter(Alfalfa_early_log2FoldChange>3 & Alfalfa_late_log2FoldChange>3)

kable(OTUs) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "400px")
```

# Compost Responders   

## Early
```{r, message = F, warning = F}
## Early   
comp.early <- subset_samples(inc.physeq, Treatment_Response %in% c("Compost_early", "Reference_early")) %>%
  filter_taxa(function(x) sum(x) >= 3, T) 

# Be very careful of the design formula in the who_diff_day() function
# This function also selects only LFC >= 2 and alpha 0.01 for significant and increasing otus to be returned
log.plot.early.comp <- comp.early %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("Compost_early", "Reference_early", comp.early) %>%
  log_plot("Compost OTUS in early group that are significantly changing compared to reference early")

# Save a data frame of these results
log.plot.early.comp.data <- log.plot.early.comp$data %>%
  mutate(trt = c("Compost_early"))

# print plot with viridis color 
log.plot.early.comp + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
```

## Late    
```{r, message = F, warning = F}
## Late 
comp.late <- subset_samples(inc.physeq, Treatment_Response %in% c("Compost_late", "Reference_late")) %>%
  filter_taxa(function(x) sum(x) >= 3, T)

# Make deseq and plot as above
log.plot.late.comp <- comp.late %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("Compost_late", "Reference_late", comp.late) %>%
  log_plot("Compost OTUS in late group that are significantly changing compared to reference late")

# Save a data frame of these results
log.plot.late.comp.data <- log.plot.late.comp$data %>%
  mutate(trt = c("Compost_late"))

# print plot with viridis color 
log.plot.late.comp + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
```

```{r, message = F, warning = F}
# Common to both early and late with LFC >=2
otustokeep <- intersect(log.plot.early.comp.data$OTU, log.plot.late.comp.data$OTU)

# Trim and rename variables
early_comp_OTUS <- log.plot.early.comp.data %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Compost_early_log2FoldChange = log2FoldChange) 

late_comp_OTUS <- log.plot.late.comp.data %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Compost_late_log2FoldChange = log2FoldChange)

# join early and late
all_comp <- full_join(early_comp_OTUS, late_comp_OTUS)
```

## Plot Compost OTUs with LFC >= 2 in both early and late  

This plot is showing the common OTUs with LFC > 3 while also showing the LFC of OTUs observed in only early or late, represented by the points landing below 2 on either axis. Common OTUs with LFC < 3 are left unlabeled.   

```{r, message = F, warning = F, out.width = "100%"}
p <- ggplot(all_comp,
       aes(x = Compost_early_log2FoldChange, 
           y = Compost_late_log2FoldChange,
           color = Phylum,
           label = OTU)) +
  geom_miss_point() +
  geom_text_repel(aes(label=ifelse(Compost_early_log2FoldChange>3 & Compost_late_log2FoldChange>3,as.character(OTU),'')),hjust=0,vjust=0) 

p + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
# Save this list of compcompa LFC > 2 for more plots
saveRDS(all_comp, file = "data/RDS/LFC_comp_OTUs_June19.RDS")
```

## Plot a table of the Compost OTUs with taxonomy   

These OTUs had response greater than 3 in both early and late compcompa   

```{r, message = F, warning = F}
OTUs <- all_comp %>%
  filter(Compost_early_log2FoldChange>3 & Compost_late_log2FoldChange>3)

kable(OTUs) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "400px")
```

# Mix Responders   

## Early
```{r, message = F, warning = F}
## Early   
mix.early <- subset_samples(inc.physeq, Treatment_Response %in% c("Mix_early", "Reference_early")) %>%
  filter_taxa(function(x) sum(x) >= 3, T) 

# Be very careful of the design formula in the who_diff_day() function
# This function also selects only LFC >= 2 and alpha 0.01 for significant and increasing otus to be returned
log.plot.early.mix <- mix.early %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("Mix_early", "Reference_early", mix.early) %>%
  log_plot("Mix OTUS in early group that are significantly changing mixared to reference early")

# Save a data frame of these results
log.plot.early.mix.data <- log.plot.early.mix$data %>%
  mutate(trt = c("Mix_early"))

# print plot with viridis color 
log.plot.early.mix + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
```

## Late    
```{r, message = F, warning = F}
## Late 
mix.late <- subset_samples(inc.physeq, Treatment_Response %in% c("Mix_late", "Reference_late")) %>%
  filter_taxa(function(x) sum(x) >= 3, T)

# Make deseq and plot as above
log.plot.late.mix <- mix.late %>%
  phyloseq_to_deseq2( ~ Treatment_Response) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("Mix_late", "Reference_late", mix.late) %>%
  log_plot("Mix OTUS in late group that are significantly changing mixared to reference late")

# Save a data frame of these results
log.plot.late.mix.data <- log.plot.late.mix$data %>%
  mutate(trt = c("Mix_late"))

# print plot with viridis color 
log.plot.late.mix + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
```

```{r, message = F, warning = F}
# Common to both early and late with LFC >=2
otustokeep <- intersect(log.plot.early.mix.data$OTU, log.plot.late.mix.data$OTU)

# Trim and rename variables
early_mix_OTUS <- log.plot.early.mix.data %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Mix_early_log2FoldChange = log2FoldChange) 

late_mix_OTUS <- log.plot.late.mix.data %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Mix_late_log2FoldChange = log2FoldChange)

# join early and late
all_mix <- full_join(early_mix_OTUS, late_mix_OTUS)
```

## Plot Mix OTUs with LFC >= 2 in both early and late  

This plot is showing the common OTUs with LFC > 3 while also showing the LFC of OTUs observed in only early or late, represented by the points landing below 2 on either axis. Common OTUs with LFC < 3 are left unlabeled.   

```{r, message = F, warning = F, out.width = "100%"}
p <- ggplot(all_mix,
       aes(x = Mix_early_log2FoldChange, 
           y = Mix_late_log2FoldChange,
           color = Phylum,
           label = OTU)) +
  geom_miss_point() +
  geom_text_repel(aes(label=ifelse(Mix_early_log2FoldChange>3 & Mix_late_log2FoldChange>3,as.character(OTU),'')),hjust=0,vjust=0) 

p + scale_colour_viridis_d(option = "plasma") +
  theme_dark()
# Save this list of mixmixa LFC > 2 for more plots
saveRDS(all_mix, file = "data/RDS/LFC_mix_OTUs_June19.RDS")
```

## Plot a table of the Mix OTUs with taxonomy   

These OTUs had response greater than 3 in both early and late mixmixa   

```{r, message = F, warning = F}
OTUs <- all_mix %>%
  filter(Mix_early_log2FoldChange>3 & Mix_late_log2FoldChange>3)

kable(OTUs) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "400px")
```