---
title: "Incubation Results"
author: "Jared"
date: "February 13, 2019"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
```
# Setup
Load needed libraries
```{r, cache=TRUE}
library(tidyverse)
library(phyloseq)
library(vegan)
library(ggplot2)
library(ggtree)
library(gplots)
library(nlme)
library(emmeans)
library(ggpubr)
library(agricolae)
library(broom)
library(xtable)
```
# Results
## Soil characteristics
### Nitrogen Mineralization
```{r, cache=TRUE}
inc.raw.physeq <- readRDS("data/RDS/incubation_physeq_Aug18.RDS")

inc.physeq <- subset_samples(inc.raw.physeq, day %in% c("7",
                                                        "14",
                                                        "21",
                                                        "35",
                                                        "49",
                                                        "97"))

#Rename treatments to more informative titles
data <- data.frame(sample_data(inc.physeq)) %>%
  mutate(treatment = recode(treatment,
                            'Control' = 'Reference',
                            'CompAlfa' = 'Mix')) %>%
  mutate(C_N = C_flash / N_flash, Inorganic_N = NH3 + NO3) %>%
  mutate(TreatmentAndDay = paste(treatment, day))
data$treatment <- relevel(data$treatment, ref = "Reference") 
data$day <- as.factor(data$day)
rownames(data) <- data$i_id
sample_data(inc.physeq) <- data
sample_data(inc.physeq)$day <- as.factor(sample_data(inc.physeq)$day)

inc.model.data <- lme(Inorganic_N~treatment * day, random=~1|replication 
                      , data = data
                      , weights = varIdent(form= ~1|day*treatment)
                      , control = lmeControl(opt = "optim", msVerbose = TRUE))

em <- emmeans(inc.model.data, c("day", "treatment"), data = data)

sum_em <- summary(em)

theme_set(theme_bw())
p <- ggplot(data = data, aes(x = day, y = Inorganic_N     )) +
  geom_point(aes(colour = treatment), size = 4) +
  stat_summary(aes(group = treatment), fun.y = mean,  geom = "line", size = 2, colour = "steelblue") +
  geom_pointrange(size = 1, pch = 1, data = sum_em, aes(x = day, y = emmean, ymin = lower.CL, ymax = upper.CL, group = treatment)) +
  xlab("Day") +
  ylab("Inorganic Nitrogen") +
  facet_wrap(~treatment) +
  theme(
    axis.title.y=element_text(colour = "black", size = 17, hjust = 0.5, margin=margin(0,12,0,0)),
    axis.title.x=element_text(colour = "black", size = 17),
    axis.text.x=element_text(colour = "black", size=15),
    axis.text.y=element_text(colour = "black", size=15),
    legend.position="none",
    legend.text=element_text(size=12.5),
    legend.key=element_blank(),
    plot.title = element_text(face = "bold"),
    strip.text.x=element_text(size=15)
  )
tiff("Figures/inorganic_N_plot.tif",height=5,width=5,units='in',res=300)
p
dev.off()
p


em2 <- emmeans(inc.model.data, c("treatment", "day"), data = data)

lambdas <- list(
  "Alfalfa - Reference" = c(-1, 1, rep(0, 24 - 2))
  , "Mix - Reference" = c(-1, 0, 1, rep(0, 24 - 3))
  , "Compost - Reference" = c(-1, 0, 0, 1, rep(0, 24 - 4))
  , "Alfalfa - Reference" = c(rep(0, 4), -1, 1, rep(0, 24 - 6))
  , "Mix - Reference" = c(rep(0, 4), -1, 0, 1, rep(0, 24 - 7))
  , "Compost - Reference" = c(rep(0, 4), -1, 0, 0, 1, rep(0, 24 - 8))
  , "Alfalfa - Reference" = c(rep(0, 8), -1, 1, rep(0, 24 - 10))
  , "Mix - Reference" = c(rep(0, 8), -1, 0, 1, rep(0, 24 - 11))
  , "Compost - Reference" = c(rep(0, 8), -1, 0, 0, 1, rep(0, 24 - 12))
  , "Alfalfa - Reference" = c(rep(0, 12), -1, 1, rep(0, 24 - 14))
  , "Mix - Reference" = c(rep(0, 12), -1, 0, 1, rep(0, 24 - 15))
  , "Compost - Reference" = c(rep(0, 12), -1, 0, 0, 1, rep(0, 24 - 16))
  , "Alfalfa - Reference" = c(rep(0, 16), -1, 1, rep(0, 24 - 18))
  , "Mix - Reference" = c(rep(0, 16), -1, 0, 1, rep(0, 24 - 19))
  , "Compost - Reference" = c(rep(0, 16), -1, 0, 0, 1, rep(0, 24 - 20))
  , "Alfalfa - Reference" = c(rep(0, 20), -1, 1, rep(0, 24 - 22))
  , "Mix - Reference" = c(rep(0, 20), -1, 0, 1, rep(0, 24 - 23))
  , "Compost - Reference" = c(rep(0, 20), -1, 0, 0, 1)
)

sum_em2 <- summary(contrast(em2, lambdas), infer = c(TRUE, TRUE), adjust = "mvt")

sum_em2$day <- factor(rep(c(7, 14, 21, 35, 49, 97), each = 3))

theme_set(theme_bw())
p <- ggplot(data = sum_em2, aes(x = day, y = estimate)) +
  geom_pointrange(aes(ymin = lower.CL, ymax = upper.CL, group = contrast, colour = contrast), size = 0.7) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  xlab("Day") +
  ylab(paste('Inorganic nitrogen \n difference from reference')) +
  scale_y_continuous(breaks = seq(-10, 800, 25)) +
  facet_wrap(~contrast) +
  theme(
    axis.title.y=element_text(colour = "black", size = 17, hjust = 0.5, margin=margin(0,12,0,0)),
    axis.title.x=element_text(colour = "black", size = 17),
    axis.text.x=element_text(colour = "black", size=15),
    axis.text.y=element_text(colour = "black", size=15),
    legend.position="none",
    legend.text=element_text(size=12.5),
    legend.key=element_blank(),
    plot.title = element_text(face = "bold"),
    strip.text.x=element_text(size=15)
  )
tiff("Figures/inorganic_N_plot_diff.tif",height=5,width=8,units='in',res=300)
p
dev.off()
p
```
#### Day 7  
```{r, cache=TRUE}
data <- data %>%
  filter(day == 7) 

g <-ggboxplot(data = data,x = "treatment"
              , y = "Inorganic_N", color = "treatment"
              , legend = "none") +
  ylab("Inorganic Nitrogen") +
  xlab("Treatment") +
  ggtitle("Day 7") +
  rotate_x_text(angle = 45) +
  ylim(0, max(data$Inorganic_N) + 5) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test", ref.group = "Reference", paired = TRUE) +
  stat_compare_means(method = "anova", label.y = max(data$Inorganic_N))
g
# linear model
lm <- lm(Inorganic_N ~ treatment, data = data)
lm.summary <- summary(lm)
lm.summary
# ANOVA
av <- aov(lm)
# Results
av.summary <- summary(av)
av.summary
# T.test x2
tukey <- TukeyHSD(av, conf.level = 0.95)
# more info from HSD.test from agricole package
tukey.1 <- HSD.test(av, trt = "treatment")
tukey.1
# latex tables
tukey.table.letters <- xtable(tukey.1$groups)
anova.table <- xtable(summary(av))
tukey.table <- xtable(tidy(tukey))
plot(lm)
```

### Microbial Biomass
### pH, water content, etc. 
## Alpha Diversity
```{r, cache=TRUE}
# This object has day 0 and amendement samples removed and is rarefied to 6000
physeq <- readRDS("data/IncPhyseqRareClusteredTree")
plot_richness(physeq)
shannon <- plot_richness(physeq, measures = "Shannon")
shannon.df <- shannon$data
# load summary function
source("Functions/summarySE.R")
summary.shannon.df <- summarySE(shannon.df, measurevar = "value", groupvars = c("treatment", "day"))
pd <- position_dodge(0.2) # move them .05 to the left and right
ggplot(summary.shannon.df, aes(x=day, y=value, colour=treatment)) + 
    geom_errorbar(aes(ymin=value-se, ymax=value+se), width=.1, position=pd) +
    geom_point(position=pd) + 
    ggtitle("Shannon diversity")
```

## Beta Diversity
Now let's plot an PCA and NMDS ordination of the weighted unifrac distances, which uses the phylogenetic tree and considers the phylogenetic relationship between OTUs when calculating the distance matrix

### PCA
```{r, cache=TRUE}
tree <- read.tree("data/tree.nwk")
physeq <- readRDS("data/RDS/IncPhyseqRareClusteredTree")
physeq.raw <- readRDS("data/RDS/incubation_physeq_Aug18.RDS")
physeq.raw <- merge_phyloseq(physeq.raw, tree)

PCoA <- ordinate(physeq, "PCoA", "wunifrac")
PCoA.raw <- ordinate(physeq.raw, "PCoA", "wunifrac")

day.PCoA.wunifrac <- plot_ordination(physeq, PCoA, color = "day") + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.2, aes(fill = day))
trt.PCoA.wunifrac <- plot_ordination(physeq, PCoA, color = "treatment") + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.2, aes(fill = treatment))
raw.day.PCoA.wunifrac <- plot_ordination(physeq.raw, PCoA.raw, color = "day") + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.2, aes(fill = day))
raw.trt.PCoA.wunifrac <- plot_ordination(physeq.raw, PCoA.raw, color = "treatment") + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.2, aes(fill = treatment))

day.PCoA.wunifrac
trt.PCoA.wunifrac
raw.day.PCoA.wunifrac
raw.trt.PCoA.wunifrac

tiff("Figures/day.PCoA.wunifrac.tif",height=4,width=9,units='in',res=300)
day.PCoA.wunifrac + 
  ggtitle("PCoA of weighted Unifrac distances on days, no amendments")
dev.off()

tiff("Figures/trt.PCoA.wunifrac.tif",height=4,width=9,units='in',res=300)
trt.PCoA.wunifrac + 
  ggtitle("PCoA of weighted Unifrac distances on treatments, no amendments")
dev.off()

tiff("Figures/raw.day.PCoA.wunifrac.tif",height=4,width=9,units='in',res=300)
raw.day.PCoA.wunifrac + 
  ggtitle("PCoA of weighted Unifrac distances on days")
dev.off()

tiff("Figures/raw.trt.PCoA.wunifrac.tif",height=4,width=9,units='in',res=300)
day.PCoA.wunifrac + 
  ggtitle("PCoA of weighted Unifrac distances on treatments")
dev.off()
```

### NMDS
```{r, cache=TRUE}
NMDS <- ordinate(physeq, "NMDS", "wunifrac")
plot_ordination(physeq, NMDS, color = "day") + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.2, aes(fill = day))
plot_ordination(physeq, NMDS, color = "treatment") + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.2, aes(fill = treatment))
```

## Aliens   
Table with # of aliens for each treatment and what period of the incubation they were detected, early/late or throughout. 

The lists of alien OTUs detected in amended microcosms was generated with <link>, we can load those lists along with the phyloseq object with clustering information from <link>:
```{r, cache=TRUE}
alf.aliens <- readRDS("data/alf.aliens.rds")
comp.aliens <- readRDS("data/comp.aliens.rds")
mix.aliens <- readRDS("data/mix.aliens.rds")
# Read in the phylsoeq object from the clustering script
```
```{r, cache=TRUE}
alf <- prune_samples(sample_data(physeq)$treatment %in% c("Alfalfa"), physeq) %>%
  filter_taxa(function(x) sum(x) > 1, T)
early.alf <- prune_samples(sample_data(alf)$response.group %in% c("early"), alf) %>%
  filter_taxa(function(x) sum(x) > 1, T)
late.alf <- prune_samples(sample_data(alf)$response.group %in% c("late"), alf) %>%
  filter_taxa(function(x) sum(x) > 1, T)

number.aliens.early.alf <- prune_taxa(as.character(alf.aliens), early.alf) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

number.aliens.late.alf <- prune_taxa(as.character(alf.aliens), late.alf) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

number.aliens.total.alf <- prune_taxa(as.character(alf.aliens), alf) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

# print("Number of alfalfa alien OTUs detected in all alfalfa")
# number.aliens.total.alf
# print("Total number of OTUs detected in all alfalfa")
# nrow(otu_table(alf))
# print("Number of alfalfa alien OTUs detected in early alfalfa")
# number.aliens.early.alf
# print("Total number of OTUs detected in early alfalfa")
# nrow(otu_table(early.alf))
# print("Number of alfalfa alien OTUs detected in late alfalfa")
# number.aliens.late.alf
# print("Total number of OTUs detected in late alfalfa")
# nrow(otu_table(late.alf))
```
```{r, cache=TRUE}
comp <- prune_samples(sample_data(physeq)$treatment %in% c("Compost"), physeq) %>%
  filter_taxa(function(x) sum(x) > 1, T)
early.comp <- prune_samples(sample_data(comp)$response.group %in% c("early"), comp) %>%
  filter_taxa(function(x) sum(x) > 1, T)
late.comp <- prune_samples(sample_data(comp)$response.group %in% c("late"), comp) %>%
  filter_taxa(function(x) sum(x) > 1, T)

number.aliens.early.comp <- prune_taxa(as.character(comp.aliens), early.comp) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

number.aliens.late.comp <- prune_taxa(as.character(comp.aliens), late.comp) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

number.aliens.total.comp <- prune_taxa(as.character(comp.aliens), comp) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

# print("Number of Compost alien OTUs detected in all Compost")
# number.aliens.total.comp
# print("Total number of OTUs detected in all Compost")
# nrow(otu_table(comp))
# print("Number of Compost alien OTUs detected in early Compost")
# number.aliens.early.comp
# print("Total number of OTUs detected in early Compost")
# nrow(otu_table(early.comp))
# print("Number of Compost alien OTUs detected in late Compost")
# number.aliens.late.comp
# print("Total number of OTUs detected in late Compost")
# nrow(otu_table(late.comp))
```
```{r, cache=TRUE}
mix <- prune_samples(sample_data(physeq)$treatment %in% c("Mix"), physeq) %>%
  filter_taxa(function(x) sum(x) > 1, T)
early.mix <- prune_samples(sample_data(mix)$response.group %in% c("early"), mix) %>%
  filter_taxa(function(x) sum(x) > 1, T)
late.mix <- prune_samples(sample_data(mix)$response.group %in% c("late"), mix) %>%
  filter_taxa(function(x) sum(x) > 1, T)

number.aliens.early.mix <- prune_taxa(as.character(mix.aliens), early.mix) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

number.aliens.late.mix <- prune_taxa(as.character(mix.aliens), late.mix) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

number.aliens.total.mix <- prune_taxa(as.character(mix.aliens), mix) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

# print("Number of Mix alien OTUs detected in all Mix")
# number.aliens.total.mix
# print("Total number of OTUs detected in all Mix")
# nrow(otu_table(mix))
# print("Number of Mix alien OTUs detected in early Mix")
# number.aliens.early.mix
# print("Total number of OTUs detected in early Mix")
# nrow(otu_table(early.mix))
# print("Number of Mix alien OTUs detected in late Mix")
# number.aliens.late.mix
# print("Total number of OTUs detected in late Mix")
# nrow(otu_table(late.mix))
```
```{r, cache=TRUE}
library(knitr)
a <- c("Alfalfa", number.aliens.total.alf, number.aliens.early.alf, number.aliens.late.alf)
b <- c("Compost", number.aliens.total.comp, number.aliens.early.comp, number.aliens.late.comp)
c <- c("Mix", number.aliens.total.mix, number.aliens.early.mix, number.aliens.late.mix)
x <- as.data.frame(rbind(a,b,c))
colnames(x) <- c("Treatment", "Throughout", "Eary", "Late")
rownames(x) <- NULL
kable(x, caption = "Number of OTUs considered aliens detected in each treatment")
```

## Alien Heatmaps
Function to get a list of OTUs from a sample type
```{r, cache=TRUE}
GetOTUs <- function(physeq, samples) {
    prune_samples(sample_data(physeq)$treatment %in% c(samples), physeq) %>%
    filter_taxa(function(x) sum(x) > 1, T) %>%
    tax_table() %>%
    row.names()
}
```

Using function to generate list of OTUs from all starting soil, incubated soils and amendments
```{r, cache=TRUE}
physeq.raw <- readRDS("data/RDS/incubation_physeq_Aug18.RDS")
# Incubated samples:
Alfalfa.otus <- GetOTUs(physeq.raw, c("Alfalfa"))
Compost.otus <- GetOTUs(physeq.raw, c("Compost"))
CompAlfa.otus <- GetOTUs(physeq.raw, c("CompAlfa"))
Control.otus <- GetOTUs(physeq.raw, c("Control"))
# Amendment samples:
AlfalfaAmend.otus <- GetOTUs(physeq.raw, c("AlfalfaAmend"))
CompostAmend.otus <- GetOTUs(physeq.raw, c("CompostAmend"))
# Soil sample:
AlfalfaSoil.otus <- GetOTUs(physeq.raw, c("AlfalfaSoil"))
```

```{r, warning=F, cache = TRUE}
GetAlienHeatMap <- function(physeq, control_otus, alien_otus, recieving_otus, samples){
  otus <- list(alien_otus, control_otus) 
  print("Looking for aliens between amendment and control soil")
  venn <- venn(otus)
  alf.aliens <- attr(venn, "intersections")$A
  aliens <- list(alf.aliens, recieving_otus)
  print("Detecting aliens in amended soil")
  aliens.venn <- venn(aliens)
  aliens.detected <- attr(aliens.venn,"intersections")$`A:B`
  incubated <- prune_samples(sample_data(physeq)$treatment %in% c(samples), physeq) %>%
    filter_taxa(function(x) sum(x) > 5, T) %>%
    transform_sample_counts(function(x) x / sum(x)) 
  incubated.aliens <- prune_taxa(aliens.detected, incubated)
  sample.order <- as.data.frame(sample_data(incubated.aliens)) %>%
    arrange(day, replication) %>%
    select(i_id) %>%
    remove_rownames()   
  alien.heatmap <- plot_heatmap(incubated.aliens, sample.label = "day", taxa.order= "Phylum", taxa.label = "Genus",
                                  sample.order = as.character(sample.order$i_id), 
                                  low = "#66CCFF", high = "#000033", na.value = "white")
  alien.heatmap
}
```

```{r, warning=FALSE, cache=TRUE}
alf.heatmap <- GetAlienHeatMap(physeq, Control.otus, AlfalfaAmend.otus, Alfalfa.otus, c("Alfalfa"))
alf.heatmap
```
```{r, warning=FALSE, cache=T}
comp.heatmap <- GetAlienHeatMap(physeq, Control.otus, CompostAmend.otus, Compost.otus, c("Compost"))
comp.heatmap
```

I didn't extract DNA from a mixed sample of compost and alfalfa, to get the list of potential aliens, we will combine the two lists from alfalfa and compost.

```{r, warning=FALSE, cache=T}
mix <- c(AlfalfaAmend.otus, CompostAmend.otus)
mix.heatmap <- GetAlienHeatMap(physeq, Control.otus, mix, CompAlfa.otus, c("Mix"))
mix.heatmap
```

The relative abundance of the alien OTUs through the incubation is relatively low, but different between treatments. Could low abundance OTUs be drivers in this situation? We should compare the relative abundance of the dominant OTUs from each treatment.

Let's make another heatmap, this one will have a list of OTUs from a treatment that is composed of the top ten OTUs by relative abundance from each day. 

```{r, warning=FALSE, cache=T}
Day_top10 <- function(physeq, trt, days){
  trt <- prune_samples(sample_data(physeq)$treatment %in% c(trt), physeq)
  
  d0 <- subset_samples(trt, day == days[1]) 
  l0 <- names(sort(taxa_sums(d0), TRUE)[1:10])
  
  d7 <- subset_samples(trt, day == days[2]) 
  l7 <- names(sort(taxa_sums(d7), TRUE)[1:10])   
  
  d14 <- subset_samples(trt, day == days[3]) 
  l14 <- names(sort(taxa_sums(d14), TRUE)[1:10])   
  
  d21 <- subset_samples(trt, day == days[4]) 
  l21 <- names(sort(taxa_sums(d21), TRUE)[1:10])
  
  d35 <- subset_samples(trt, day == days[5]) 
  l35 <- names(sort(taxa_sums(d35), TRUE)[1:10])
  
  d49 <- subset_samples(trt, day == days[6]) 
  l49 <- names(sort(taxa_sums(d49), TRUE)[1:10])
  
  d97 <- subset_samples(trt, day == days[7])
  l97 <- names(sort(taxa_sums(d97), TRUE)[1:10])
  list <- c(l0, l7, l14, l21, l35, l49, l97)
  list
  phy <- prune_taxa(list, trt) %>%
      filter_taxa(function(x) sum(x) > 5, T) %>%
      transform_sample_counts(function(x) x / sum(x))
  sample.order <- as.data.frame(sample_data(phy)) %>%
    arrange(day, replication) %>%
    select(i_id) %>%
    remove_rownames()   
  heatmap <- plot_heatmap(phy, sample.label = "day", taxa.order= "Phylum", taxa.label = "Genus",
                                  sample.order = as.character(sample.order$i_id), 
                                  low = "#66CCFF", high = "#000033", na.value = "white")
  heatmap
}
```

Not rarefied for these heatmaps:
```{r, warning=FALSE, cache=T}
days <- c("0", "7", "14", "21", "35", "49", "97")
alf <- Day_top10(physeq.raw, c("Alfalfa"), days)
alf
```

Compost:
```{r, warning=FALSE, cache=T}
comp <- Day_top10(physeq.raw, c("Compost"), days)
comp
```

Mix:
```{r, warning=FALSE, cache=T}
mix <- Day_top10(physeq.raw, c("CompAlfa"), days)
mix
```

What are the common top OTUs from the three treatments? By treatment, the total number of OTUs made up of each treatment + day's top 10 was different. `levels(comp$data$OTU)` will show you the unique OTUs. 
```{r, cache=T}
venn(list("Alfalfa" = levels(alf$data$OTU), "Compost" = levels(comp$data$OTU), "Mix" = levels(mix$data$OTU)))
```


