---
title: "Incubation Results"
author: "Jared"
date: "February 13, 2019"
output: 
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
```
# Setup
Load needed libraries
```{r, cache=TRUE}
library(tidyverse)
library(phyloseq)
library(vegan)
library(ggplot2)
library(ggtree)
```
# Results

## Alpha Diversity
```{r}
# This object has day 0 and amendement samples removed and is rarefied to 6000
physeq <- readRDS("data/IncPhyseqRareClusteredTree")
plot_richness(physeq)
shannon <- plot_richness(physeq, measures = "Shannon")
shannon.df <- shannon$data
# load summary function
source("Functions/summarySE.R")
summary.shannon.df <- summarySE(shannon.df, measurevar = "value", groupvars = c("treatment", "day"))
pd <- position_dodge(0.2) # move them .05 to the left and right
ggplot(summary.shannon.df, aes(x=day, y=value, colour=treatment)) + 
    geom_errorbar(aes(ymin=value-se, ymax=value+se), width=.1, position=pd) +
    geom_point(position=pd)
```

## Beta Diversity
Now let's plot an PCA and NMDS ordination of the weighted unifrac distances, which uses the phylogenetic tree and considers the phylogenetic relationship between OTUs when calculating the distance matrix

### PCA
```{r}
PCoA <- ordinate(physeq, "PCoA", "wunifrac")
plot_ordination(physeq, PCoA, color = "day") + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.2, aes(fill = day))
plot_ordination(physeq, PCoA, color = "treatment") + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.2, aes(fill = treatment))
```

### NMDS
```{r, }
NMDS <- ordinate(physeq, "NMDS", "wunifrac")
plot_ordination(physeq, NMDS, color = "day") + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.2, aes(fill = day))
plot_ordination(physeq, NMDS, color = "treatment") + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.2, aes(fill = treatment))
```

## Aliens   
Table with # of aliens for each treatment and what period of the incubation they were detected, early/late or throughout. 

The lists of alien OTUs detected in amended microcosms was generated with <link>, we can load those lists along with the phyloseq object with clustering information from <link>:
```{r}
alf.aliens <- readRDS("data/alf.aliens.rds")
comp.aliens <- readRDS("data/comp.aliens.rds")
mix.aliens <- readRDS("data/mix.aliens.rds")
# Read in the phylsoeq object from the clustering script
```
```{r}
alf <- prune_samples(sample_data(physeq)$treatment %in% c("Alfalfa"), physeq) %>%
  filter_taxa(function(x) sum(x) > 1, T)
early.alf <- prune_samples(sample_data(alf)$response.group %in% c("early"), alf) %>%
  filter_taxa(function(x) sum(x) > 1, T)
late.alf <- prune_samples(sample_data(alf)$response.group %in% c("late"), alf) %>%
  filter_taxa(function(x) sum(x) > 1, T)

number.aliens.early.alf <- prune_taxa(as.character(alf.aliens), early.alf) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

number.aliens.late.alf <- prune_taxa(as.character(alf.aliens), late.alf) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

number.aliens.total.alf <- prune_taxa(as.character(alf.aliens), alf) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

# print("Number of alfalfa alien OTUs detected in all alfalfa")
# number.aliens.total.alf
# print("Total number of OTUs detected in all alfalfa")
# nrow(otu_table(alf))
# print("Number of alfalfa alien OTUs detected in early alfalfa")
# number.aliens.early.alf
# print("Total number of OTUs detected in early alfalfa")
# nrow(otu_table(early.alf))
# print("Number of alfalfa alien OTUs detected in late alfalfa")
# number.aliens.late.alf
# print("Total number of OTUs detected in late alfalfa")
# nrow(otu_table(late.alf))
```
```{r}
comp <- prune_samples(sample_data(physeq)$treatment %in% c("Compost"), physeq) %>%
  filter_taxa(function(x) sum(x) > 1, T)
early.comp <- prune_samples(sample_data(comp)$response.group %in% c("early"), comp) %>%
  filter_taxa(function(x) sum(x) > 1, T)
late.comp <- prune_samples(sample_data(comp)$response.group %in% c("late"), comp) %>%
  filter_taxa(function(x) sum(x) > 1, T)

number.aliens.early.comp <- prune_taxa(as.character(comp.aliens), early.comp) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

number.aliens.late.comp <- prune_taxa(as.character(comp.aliens), late.comp) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

number.aliens.total.comp <- prune_taxa(as.character(comp.aliens), comp) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

# print("Number of Compost alien OTUs detected in all Compost")
# number.aliens.total.comp
# print("Total number of OTUs detected in all Compost")
# nrow(otu_table(comp))
# print("Number of Compost alien OTUs detected in early Compost")
# number.aliens.early.comp
# print("Total number of OTUs detected in early Compost")
# nrow(otu_table(early.comp))
# print("Number of Compost alien OTUs detected in late Compost")
# number.aliens.late.comp
# print("Total number of OTUs detected in late Compost")
# nrow(otu_table(late.comp))
```
```{r}
mix <- prune_samples(sample_data(physeq)$treatment %in% c("Mix"), physeq) %>%
  filter_taxa(function(x) sum(x) > 1, T)
early.mix <- prune_samples(sample_data(mix)$response.group %in% c("early"), mix) %>%
  filter_taxa(function(x) sum(x) > 1, T)
late.mix <- prune_samples(sample_data(mix)$response.group %in% c("late"), mix) %>%
  filter_taxa(function(x) sum(x) > 1, T)

number.aliens.early.mix <- prune_taxa(as.character(mix.aliens), early.mix) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

number.aliens.late.mix <- prune_taxa(as.character(mix.aliens), late.mix) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

number.aliens.total.mix <- prune_taxa(as.character(mix.aliens), mix) %>%
  filter_taxa(function(x) sum(x) > 1, T) %>%
  otu_table() %>%
  nrow()

# print("Number of Mix alien OTUs detected in all Mix")
# number.aliens.total.mix
# print("Total number of OTUs detected in all Mix")
# nrow(otu_table(mix))
# print("Number of Mix alien OTUs detected in early Mix")
# number.aliens.early.mix
# print("Total number of OTUs detected in early Mix")
# nrow(otu_table(early.mix))
# print("Number of Mix alien OTUs detected in late Mix")
# number.aliens.late.mix
# print("Total number of OTUs detected in late Mix")
# nrow(otu_table(late.mix))
```
```{r}
library(knitr)
a <- c("Alfalfa", number.aliens.total.alf, number.aliens.early.alf, number.aliens.late.alf)
b <- c("Compost", number.aliens.total.comp, number.aliens.early.comp, number.aliens.late.comp)
c <- c("Mix", number.aliens.total.mix, number.aliens.early.mix, number.aliens.late.mix)
x <- as.data.frame(rbind(a,b,c))
colnames(x) <- c("Treatment", "Throughout", "Eary", "Late")
rownames(x) <- NULL
kable(x, caption = "Number of OTUs considered aliens detected in each treatment")
```

## Alien Heatmaps


