---
title: "Native Persistor OTUs in the incubation experiment"
output: html_document
---
# Native Persistor   
`Persistor: OTUs detected in day 7 soils that persist into early or early and late groups of each treatment group` 
`Native: OTUs in Reference soil for a given time point`     
---
Create a list of OTUs that make up the core taxa from reference soils. Let's create a core list for day 0, 0-97(all days), early and late groups. We will end up with four lists of OTUs for each grouping.
```{r, echo=F, warning=F, message=F}
library(phyloseq)
library(vegan)
library(tidyverse)
library(gplots)
library(DESeq2)
```

```{r, message=F, warning=F}
inc.physeq <- readRDS("../data/RDS/incubation_physeq_Aug18.RDS")

tree <- read.tree("../data/tree.nwk")

inc.physeq <- merge_phyloseq(inc.physeq, tree)

#Rename treatments to more informative titles
data <- data.frame(sample_data(inc.physeq)) %>% 
  mutate(treatment = recode(treatment,
                            'Control' = 'Reference',
                            'CompAlfa' = 'Mix')) %>% 
  mutate(C_N = C_flash / N_flash, Inorganic_N = NH3 + NO3) %>%
  mutate(TreatmentAndDay = paste(treatment, day))

rownames(data) <- data$i_id
sample_data(inc.physeq) <- data
sample_data(inc.physeq)$day <- as.factor(sample_data(inc.physeq)$day)

inc.physeq.data <- data.frame(sample_data(inc.physeq))
inc.physeq.data$response.group[inc.physeq.data$day == "0"] <-
"baseline"
inc.physeq.data$response.group[inc.physeq.data$day %in% c("7", "14", "21")] <-
"early"
inc.physeq.data$response.group[inc.physeq.data$day %in% c("35", "49", "97")] <-
"late"

inc.physeq.data <- inc.physeq.data %>% 
  mutate(Treatment_Response = paste(treatment, response.group, sep = '_'))

rownames(inc.physeq.data) <- data$i_id
sample_data(inc.physeq) <- inc.physeq.data
```

```{r, message=F}
ref_7 <- subset_samples(inc.physeq, treatment %in% c("Reference") & day %in% c("7")) %>%
  filter_taxa(function(x) sum(x) > 0, T)

otu.mean = apply(X = otu_table(ref_7), MARGIN = 1, mean)
otu.freq = rowSums(otu_table(ref_7) != 0)

t = data.frame(Frequency = otu.freq, 
               Mean = otu.mean,
               Total_Abundance = taxa_sums(ref_7), 
               tax_table(ref_7))
```
How about looking at high frequency OTUs? How many OTUs were detected in 8 out of 12 alfalfa microcosms on day 7?     
```{r, message=F}
nrow(ref_7)
high.freq <- t %>% 
  rownames_to_column("OTU") %>%
  filter(Frequency >= 8) 

nrow(high.freq)
```
Track relative abundance of these 795 high frequency day 7 Reference OTUs throughout the incubated alfalfa timeline.    
```{r, message=F, warning=F}
Alfalfa <- subset_samples(inc.physeq, treatment %in% c("Alfalfa") & response.group %in% c("early", "late"))
# Prune incubated Alfalfa samples to core from reference day 7
Alfalfa.high.freq <- prune_taxa(high.freq$OTU, Alfalfa) %>%
  filter_taxa(function(x) sum(x) > 0, T) %>%
  transform_sample_counts(function(x) x/sum(x))

myTaxa <- names(sort(taxa_sums(Alfalfa.high.freq), decreasing = T)[1:20])

Alfalfa.high.freq.20 <- prune_taxa(myTaxa, Alfalfa.high.freq) %>%
  filter_taxa(function(x) sum(x) > 0, T)

nrow(tax_table(Alfalfa.high.freq.20))
# This will make a list of samples in the correct day order
sample.order <- as.data.frame(sample_data(Alfalfa.high.freq.20)) %>%
  arrange(day, replication) %>%
  select(i_id) %>%
  remove_rownames() 

plot_heatmap(Alfalfa.high.freq.20, sample.label = "day", taxa.order = "Phylum", taxa.label = "Genus",  
                              sample.order = as.character(sample.order$i_id), 
                              low = "yellow", high = "red", na.value = "grey") 

plot_tree(Alfalfa.high.freq.20, label.tips = "Genus", ladderize = "left", justify = "left", size = "Abundance")
```

```{r, message=F, warning=F}
Compost <- subset_samples(inc.physeq, treatment %in% c("Compost") & response.group %in% c("early", "late"))
# Prune incubated Compost samples to core from reference day 7
Compost.high.freq <- prune_taxa(high.freq$OTU, Compost) %>%
  filter_taxa(function(x) sum(x) > 0, T) %>%
  transform_sample_counts(function(x) x/sum(x))

myTaxa <- names(sort(taxa_sums(Compost.high.freq), decreasing = T)[1:20])

Compost.high.freq.20 <- prune_taxa(myTaxa, Compost.high.freq) %>%
  filter_taxa(function(x) sum(x) > 0, T)

nrow(tax_table(Compost.high.freq.20))
# This will make a list of samples in the correct day order
sample.order <- as.data.frame(sample_data(Compost.high.freq.20)) %>%
  arrange(day, replication) %>%
  select(i_id) %>%
  remove_rownames() 

plot_heatmap(Compost.high.freq.20, sample.label = "day", taxa.order = "Phylum", taxa.label = "Genus",  
                              sample.order = as.character(sample.order$i_id), 
                              low = "yellow", high = "red", na.value = "grey") 

plot_tree(Compost.high.freq.20, label.tips = "Genus", ladderize = "left", justify = "left", size = "Abundance")
```

```{r, message=F, warning=F}
Mix <- subset_samples(inc.physeq, treatment %in% c("Mix") & response.group %in% c("early", "late"))
# Prune incubated Mix samples to core from reference day 7
Mix.high.freq <- prune_taxa(high.freq$OTU, Mix) %>%
  filter_taxa(function(x) sum(x) > 0, T) %>%
  transform_sample_counts(function(x) x/sum(x))

myTaxa <- names(sort(taxa_sums(Mix.high.freq), decreasing = T)[1:20])

Mix.high.freq.20 <- prune_taxa(myTaxa, Mix.high.freq) %>%
  filter_taxa(function(x) sum(x) > 0, T)

nrow(tax_table(Mix.high.freq.20))
# This will make a list of samples in the correct day order
sample.order <- as.data.frame(sample_data(Mix.high.freq.20)) %>%
  arrange(day, replication) %>%
  select(i_id) %>%
  remove_rownames() 

plot_heatmap(Mix.high.freq.20, sample.label = "day", taxa.order = "Phylum", taxa.label = "Genus",  
                              sample.order = as.character(sample.order$i_id), 
                              low = "yellow", high = "red", na.value = "grey") 

plot_tree(Mix.high.freq.20, label.tips = "Genus", ladderize = "left", justify = "left", size = "Abundance")
```

